\section{Evaluation}\label{sec:results}

\subsection{Performance}

\tblBenchmarks

\autoref{tbl:benchmarks} compares the performance of CraterLake, CPU, and F1+
on deep and shallow benchmarks. It shows execution times and CraterLake's
speedups over the CPU and F1+.

CraterLake achieves very large speedups over the CPU implementations, ranging
from 2,900$\times$ to 8,600$\times$. Speedups are similar across deep and
shallow benchmarks, showing that CraterLake provides robust gains across FHE
program types.

CraterLake achieves large speedups over F1+ on deep benchmarks, outperforming
it by gmean 11.2$\times$ and up to 18.6$\times$. By contrast, CraterLake and F1+
achieve comparable speeds on shallow benchmarks, with a gmean speedup of only
1.34$\times$. These stark differences demonstrate that \emph{prior FHE
accelerators are efficient only on shallow computations, and falter on deep
ones due to the limitations that we have discussed}.

On the MNIST shallow benchmarks, CraterLake is slightly slower than F1+ because
standard keyswitching is efficient in this range and CraterLake spends a large
fraction of~area~on~the CRB, which has low utilization at low $L$, whereas F1+
has higher NTT, add, and multiply throughput. But the high speedups on deep
programs show that most compute happens at high $L$, so CraterLake optimizes
for~the~common~case.

\subsection{Architectural Analysis}\label{sec:architecturalAnalysis}

To understand these results in more depth, we examine CraterLake's resource
utilization and power consumption.

\tblUtilization

\paragraph{Utilization:}
\autoref{tbl:utilization} reports the average utilization of FUs and main
memory bandwidth for each application. FU utilization is reported as the
fraction of cycles that FUs are consuming input values, averaged across all
FUs. It reflects \emph{issue rate}---for example, a utilization of 66\% means
that 10 out of the 15 FUs are consuming new inputs. Bandwidth utilization is
simply the fraction of cycles memory is active, e.g., 70\% utilization denotes
an average bandwidth of 700\,GB/s.

Overall, CraterLake achieves high utilization of both memory and compute,
denoting a balanced system. Thanks to the CRB, CraterLake's FU mix is balanced
across all ciphertext sizes, so FU utilization is always high unless memory
limits throughput. Some workloads, like unpacked bootstrapping, saturate on
memory bandwidth, but most others see compute utilization near or above 50\%
(near 100\% utilization is not achievable as all programs experience some
memory-bound phases).

By contrast, F1+'s utilization is much lower, especially on deep benchmarks,
where average FU utilization is 10\%, owing to its inadequate FU mix and lack
of CRB.

\figDataMovementAndPower

\paragraph{Data movement:} \autoref{fig:dataMovement}
breaks down memory traffic by keyswitch hint (KSH), inputs, and intermediate
loads and stores. We see that deep benchmarks incur a manageable amount of
intermediate traffic, while shallow benchmarks have a sufficiently low
footprint to cause no eviction of intermediates.

\paragraph{Power consumption.}
\autoref{fig:power} shows the power breakdown for CraterLake and the total
power consumption for each benchmark. The figure includes both chip and HBM
power. Power stays within a 320\,W envelope, and is higher for deep benchmarks,
primarily due to higher FU/memory utilization and higher internal CRB
utilization. FUs dominate power across benchmarks, consuming 50-80\%. This
shows the importance of CraterLake's FU energy optimizations, and demonstrates
that its architecture is far more efficient: F1 was dominated by data movement
energy, and F1+ consumes gmean 18$\times$ more energy than CraterLake on our
deep benchmarks. CraterLake's efficiency and performance benefits yield a gmean
201$\times$ improvement in performance per Joule over F1+.

\subsection{Sensitivity Studies}
\label{sec:sensitivity}

\figGmeanVsStorage

\paragraph{On-chip storage:}
\autoref{fig:gmeanVsStorage} shows CraterLake's performance as register file
grows from 100 to 350\,MB. Each line shows the performance of a different
application, \emph{normalized to its performance at the default size, 256\,MB}.
While shallow benchmarks are insensitive to storage size, most deep benchmarks
suffer severely from a smaller register file, incurring slowdowns of up to
5.5$\times$. This shows that CraterLake's large on-chip storage is crucial for
deep benchmarks. Finally, adding more on-chip storage leads to diminishing
returns.

\tblPerformanceBreakdown

\paragraph{Effect of CraterLake's features:}
\autoref{tbl:performanceBreakdown} shows the impact of our innovations by
reporting the \emph{slowdown} of alternative configurations. \emph{KSHGen}
omits the KSHGen FU and stores full KSHs in memory. This hurts performance
noticeably, especially in deep benchmarks, by gmean 1.9$\times$.
\emph{CRB/chain} omits the CRB and the vector chaining optimizations.
Performance is even worse than for F1+, because the system becomes bottlenecked
on RF ports and CraterLake has 50\% of the NTT and 40\% of the multiply/add
throughput of F1+. \emph{Network} replaces CraterLake's fixed transpose network
and polynomial tiling design with F1+'s crossbars and residue polynomial tiling
design. Performance is up to 2$\times$ worse \emph{even though the F1+ network
is 16$\times$ larger}, because residue polynomial tiling incurs 2.4$\times$
more traffic than CraterLake's approach (\autoref{sec:comparison}).

\subsection{Performance vs.\ Target Security Level}
\label{sec:moreSecurity}

The benchmarks presented so far meet an 80-bit security level, which is often
considered
sufficient~\cite{halevi2021bootstrapping,halevi2018faster,izabachene2019practical,ji2019efficient}.
We now analyze the impact of meeting higher levels of security: 128-bit
security, a commonly used
target~\cite{albrecht:hesg18:standard,lee:2021:privacy,riazi:asplos20:heax}
that can be efficiently achieved with $N =64K$, and 200-bit security, a very
conservative target that requires using larger polynomials.

\paragraph{128-bit security:}
\autoref{tbl:securebenchmarks} shows CraterLake's performance on our benchmarks
when adjusted for 128-bit security (128-bit column), as well as the slowdown
compared to performance for 80-bit security.

\tblSecureBenchmarks

To reach 128 bits of security with a polynomial degree of $N$=64K, we bootstrap
twice as often as with 80 bits of security, i.e., target half the number of
usable levels after bootstrapping. This allows using boosted keyswitching
variants with a relatively small number of digits: we use 1-digit keyswitching
for $L< 32$, 2-digit keyswitching for $32 \leq L< 43$, and 3-digit keyswitching
for $43 \leq L$. As we bootstrap twice as often, we never go beyond $L =51$.

\autoref{tbl:securebenchmarks} (left) shows that 128-bit security adds modest
overheads: a 1.36$\times$ gmean slowdown, and a worst-case slowdown of
1.62$\times$. These overheads stem from the higher memory footprint of
multi-digit keyswitching and the more frequent bootstrapping. Importantly,
though bootstrapping is twice as frequent, slowdowns are below 2$\times$,
because both bootstrapping and useful computation happen at lower $L$ values,
lowering their cost. This is a concrete example of the tradeoff that
\autoref{fig:bootstrappingFrequency} in \autoref{sec:deepChallenges}
illustrates.

\paragraph{Beyond 128-bit security:}
Effectively supporting significantly more than 128-bit security requires using
larger polynomials. Here, we evaluate a 200-bit security target, which requires
doubling $N$ from $64K$ to $128K$. Note that this security target is very
conservative, and not used in FHE benchmarks; we use it to study performance
over a wide range of security levels.

CraterLake as evaluated so far support $N$ up to 64K natively, but larger
vectors would require multiple passes through FUs and forgo FU chaining. We
thus evaluate a CraterLake version that supports $N$ up to 128K natively. This
requires modest hardware changes, chiefly \emph{(1)} the buffers in the CRB
need to double, from 26.25$\,$MB to 52.5$\,$MB; and \emph{(2)} NTTs require an
additional butterfly stage. These changes consume 27.4\,mm$^2$ of additional
area, i.e., less than $7\%$ of chip area.

\autoref{tbl:securebenchmarks} (right) reports the performance of deep
benchmarks for 200-bit security and $N$=128K. Since $N$=128K ciphertexts have
double the slots of $N$=64K ciphertexts, we normalize performance per element.
This is because doubling $N$ allows doubling the number of slots (plaintext
elements) per ciphertext, which enables more computations to happen in
parallel. For example, consider the ResNet benchmark: starting from the
original benchmark, which uses $N$=64K ciphertexts and performs one inference
at a time, it is easy to construct a ResNet benchmark that uses $N =128K$ and
performs two inferences in parallel. In addition to doubling $N$, CraterLake
achieves a 200-bit security by using higher-digit keyswitching.

\autoref{tbl:securebenchmarks} shows that the 200-bit security target imposes
additional overheads, incurring a 2.6$\times$ gmean slowdown over 80-bit
security, and a worst-case slowdown of 4.35$\times$. Most of these slowdown are
caused by the fact that using $N =128K$ doubles the footprint of ciphertexts and
KSHs over $N =64K$. This limits reuse opportunities and adds off-chip traffic.
While doubling the register file (to 512\,MB) would erase most of these
overheads, it would add significant area.
