This work was conducted in collaboration with Nikola Samardzic, Alex Krastev, Nicholas Genise, Prof. Srini Devadas, Karim Eldefrawy, Prof. Ron Dreslinski, Prof. Christopher Peikert, and my research advisor Prof. Daniel Sanchez. Much of this thesis is adapted from a joinly written paper. This work would not have been possible without all of their contributions, and I have learned a lot working with them. Fully Homomorphic Encryption is not a very accessible research topic, and I am grateful to my collaborators for always taking the time to explain complex mathematical notions in simple terms that I can understand.

I am especially grateful to my advisor, Prof. Daniel Sanchez. It is very easy to be impressed by Daniel's technical ability and encyclopedic knowledge of computer architecture. However, I am most thankful for Daniel's infinite patience and constant encouragement. Working in isolation during a global pandemic has been tough, but Daniel's consistent support has made a difficult situation much easier. Whenever I have felt lost, Daniel has always had time to meet with me and discuss whatever is on my mind, be it high level research directions or tiny specific software bugs.

I am also extremely grateful to my research partner, Nikola Samardzic. Many times, when I have felt very stuck on a particular problem, Nikola has calmly stepped up and engineered a brilliant solution. Though I have never actually met Nikola in person, I feel that working on this project has brought us together, and I consider Nikola to be a dear friend.

I would also like to thank all the members of our research group. My officemates, Quan Nguyen, Yifan Yang, and Victor Ying have always been willing to help me out with both technical and non-technical problems. On top of that, they have always put up with my jokes and unrelated discussion topics. My other groupmates, Maleen Abeydeera, Prof. Joel Emer, Hyun Ryong (Ryan) Lee, and Guowei Zhang have always provided me with fascinating research discussions. I have learned so much from all of them, and our group hangouts have been a great source of joy during a relatively bleak time period.

I would like to thank my academic advisor, Prof. Vivienne Sze, who has always been supportive and given me good advice on classes and research. 

I would also like to thank my undergraduate-research advisor at CMU, Prof. Nathan Beckmann, for getting me started in computer architecture research and always giving me great advice. Without him, I would not be where I am today.

Last but certainly not least, I would like to thank my parents. My parents have always provided me with unconditional support in every endeavour. They have always encouraged me to seek to better myself, take risks, and pursue my goals. And when I fail, they have always been there to catch me. Anything I have achieved in my life so far would not have been possible without their love and sacrifices.